---
title: "QC Plots"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r,include=FALSE}
library(tidyverse)
library(VIM)
library(ggrepel)
library(gridExtra)
library(readr)
QC.df <- read_csv("~/Projects/ExtraLong/Data/Template_Quality_AllSubjects.csv")
aseg.df <- read_delim("~/Projects/ExtraLong/Data/freesurferLongitudinal_aseg.table.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
ScanTime.df <- read_table2("~/Projects/ExtraLong/Data/qdec_table.dat")
```

The following code examines NA values across each column of the data set. It appears we have twenty individuals who have NA values for all CNR values, holes_total, and euler_total. 

```{r}
aggr(QC.df,numbers = TRUE,prop = FALSE,cex.axis = .9,oma = c(10,4,3,.5))
```

By printing out the twenty individuals who have NA values, we see that the values for euler_lh, euler_rh, holes_lh, and holes_rh are also mistaken. 

```{r}
QC.df %>% 
  filter(is.na(cnr_graycsf_lh)) %>% 
  relocate(bblid,euler_lh,euler_rh,holes_lh,holes_rh)
```

Removes individuals with any NA values from the data set. 
```{r}
QC.trim <- QC.df %>% 
  filter_all(all_vars(!is.na(.))) %>% 
  mutate(holes_lh = as.numeric(holes_lh),holes_rh = as.numeric(holes_rh),euler_lh=as.numeric(euler_lh),euler_rh=as.numeric(euler_rh))
```

Creates plots for QC values in QC.df

```{r}
QC.measures <- QC.trim %>% 
  select(-bblid,-seslabel) %>% 
  colnames() 

names.df <- data.frame(Short.Name = c("cnr_graycsf_lh","cnr_graycsf_rh","cnr_graywhite_lh","cnr_graywhite_rh","holes_lh","holes_rh","holes_total","euler_lh","euler_rh","euler_total"),Long.Name = c("CNR Gray-CSF Left Hemisphere", "CNR Gray-CSF Right Hemisphere", "CNR Gray-White Left Hemisphere", "CNR Gray-White Right Hemisphere","Holes Left Hemisphere","Holes Right Hemisphere","Holes Total","Euler Left Hemisphere","Euler Right Hemisphere","Euler Total"))
  
getQCplot <- function(Measure.of.Interest){
  
  if(str_detect(Measure.of.Interest,pattern = "holes")){
    
    Name.of.Measure <- names.df %>% 
      filter(Short.Name == Measure.of.Interest) %>% 
      pull(Long.Name)
    
    Measure.of.Interest <- quo(!!sym(Measure.of.Interest))
    
    QC.trim <- QC.trim %>% 
      mutate(bblid = as.factor(bblid)) %>% 
      mutate(bblid = fct_reorder(bblid,!!Measure.of.Interest))
    
    QC.trim.label <- QC.trim %>% 
      slice_max(order_by = !!Measure.of.Interest,n = 10)
      
    QC.trim.bblid <- QC.trim %>% 
      slice_max(order_by = !!Measure.of.Interest,n = 10) %>% 
      pull(bblid)
    
    QC.trim %>% 
      mutate(Worst10 = ifelse(bblid %in% QC.trim.bblid,"Worst 10","Other")) %>% 
      ggplot(aes(x = bblid,y = !!Measure.of.Interest)) + geom_point(aes(color = Worst10)) + scale_x_discrete(breaks = NULL,expand = expansion(add = 15)) + theme(axis.text.x = element_blank(),axis.ticks.x = element_blank(),panel.grid.major = element_blank(),panel.grid.minor = element_blank(),panel.background = element_blank(),axis.line = element_line(color = "black")) + labs(title = paste(Name.of.Measure),x = "ID #") + geom_label_repel(data = QC.trim.label,aes(label=bblid),max.overlaps = getOption("ggrepel.max.overlaps", default = Inf)) + scale_color_brewer(name = "Low Quality Templates",palette = "Dark2") 
  } else{
    
    Name.of.Measure <- names.df %>% 
      filter(Short.Name == !!Measure.of.Interest) %>% 
      pull(Long.Name)
    
    Measure.of.Interest <- quo(!!sym(Measure.of.Interest))

    QC.trim <- QC.trim %>% 
      mutate(bblid = as.factor(bblid)) %>% 
      mutate(bblid = fct_reorder(bblid,!!Measure.of.Interest))
    
    QC.trim.label <- QC.trim %>% 
      slice_min(order_by = !!Measure.of.Interest,n = 10)
      
    QC.trim.bblid <- QC.trim %>% 
      slice_min(order_by = !!Measure.of.Interest,n = 10) %>% 
      pull(bblid)
    
    QC.trim %>% 
      mutate(Worst10 = ifelse(bblid %in% QC.trim.bblid,"Worst 10","Other")) %>% 
      ggplot(aes(x = bblid,y = !!Measure.of.Interest)) + geom_point(aes(color = Worst10)) + scale_x_discrete(breaks = NULL,expand = expansion(add = 15)) + theme(axis.text.x = element_blank(),axis.ticks.x = element_blank(),panel.grid.major = element_blank(),panel.grid.minor = element_blank(),panel.background = element_blank(),axis.line = element_line(color = "black")) + labs(x = "ID #",title = paste(Name.of.Measure)) + geom_label_repel(data = QC.trim.label,aes(label=bblid),max.overlaps = getOption("ggrepel.max.overlaps", default = Inf)) + scale_color_brewer(name = "Low Quality Templates",palette = "Dark2") 
  }
}

QCplots <- map(QC.measures,getQCplot)
```

QC plots

```{r,echo = FALSE,warning=FALSE}
QCplots[[1]]
QCplots[[2]]
QCplots[[3]]
QCplots[[4]]
QCplots[[5]]
QCplots[[6]]
QCplots[[7]]
QCplots[[8]]
QCplots[[9]]
QCplots[[10]]
```

Rank for Each Template 

```{r}
# The code determines a participant's rank (low numbers indicating bad image quality) compared to other participants for each QC measure. For example, if participant #123 has the lowest euler_total value out of all participants, they receive a ranking of 1 for euler_total.Then, a total rank is calculated for each participant by adding up their ranking across all QC measures. 
N <- nrow(QC.trim)
RankByID <- QC.trim %>% 
  pivot_longer(cols = cnr_graycsf_lh:euler_total,names_to = "QC.MeasureType",values_to = "Values") %>% 
  mutate(QC.MeasureType = ifelse(str_count(QC.MeasureType,pattern = "_")>1,str_replace(QC.MeasureType,pattern = "_",replacement = ""),QC.MeasureType)) %>% 
  separate(QC.MeasureType,into = c("QC.MeasureType","Region")) %>% 
  group_by(QC.MeasureType,Region) %>% 
  mutate(rank = min_rank(Values)) %>% 
  ungroup() %>% 
  mutate(rank = ifelse(str_detect(QC.MeasureType,pattern = "holes"),N-rank+1,rank)) %>% 
  group_by(bblid) %>% 
  summarize(Total.Rank = sum(rank)) %>% 
  arrange(Total.Rank) 
head(RankByID)
```

##Examining aseg values within and across subjects

First, we will add in the scan date for each session so we can plot subjects longitudinally.
```{r}
aseg.df <- aseg.df %>% 
  rename(id = `Measure:volume`) %>% 
  mutate(id = str_remove(id,pattern = "\\.long.Template.*")) %>% 
  left_join(ScanTime.df,by = c("id" = "fsid")) %>% 
  rename(Scan.Date = `Scan-Date`) %>% 
  relocate(id,Scan.Date) %>%
  mutate(Scan.Date = str_remove(Scan.Date,pattern = "T.*")) %>% 
  mutate(Scan.Date = str_remove(Scan.Date,pattern = '"')) %>% 
  mutate(Scan.Date = as.Date(Scan.Date)) %>% 
  separate(id, into = c("Subject.ID","Session"),sep = "/") %>% 
  mutate(Subject.ID = str_remove(Subject.ID,pattern = "sub-"))
```

At first glance, the estimated total intracranial volume appears not to change across time for any individual. 

```{r}
aseg.df %>% 
  ggplot(aes(x = Scan.Date,y = EstimatedTotalIntraCranialVol,group = Subject.ID)) + geom_point(color = "lightsteelblue") + geom_line(color = "lightsteelblue") + theme_classic() + labs(x = "Scan Date",y = "Estimated Total Intracranial Volume",title = "Intracranial Volume by Subject")
```

The following data frame confirms this, as the highest variability of intracranial volume for a particular subject is 0 (i.e across scan dates estimated intracranial volume never changes).

```{r}
aseg.df %>% 
  group_by(Subject.ID) %>% 
  summarize(IntraCranialVolVariance = var(EstimatedTotalIntraCranialVol)) %>% 
  arrange(desc(IntraCranialVolVariance)) %>% 
  head()
```


















